Merging Datasets Through Deep learning

### 2018-08-29 paper2018/main.tex
ontology propaganda

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity.  We propose a novel approach to performing merges based on deep learning models.  Our approach depends on (a) creating a deep learning model that maps surface forms of an entity into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow for joins with precision of .91-.93 and recall of .70-.81. To build these models, we had to adapt techniques from metric learning due to the properties of the our domain.  We demonstrate sample selection and loss functions, which work for this domain.  We make the models available to the community for aligning people or companies across multiple datasets.

\end{abstract}

>>>>

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent
requirement for merging is joining across columns that have
different surface forms for the same entity (e.g., the name of a
person might be represented as \textit{Douglas Adams},
\textit{Adams, Douglas} or \textit{Douglas Noel Adams}).  Similarly,
ontology alignment can require recognizing distinct surface forms of
the same entity, especially when ontologies are independently
developed.  However, data management systems are currently limited
to performing merges based on string equality, or at best using
string similarity.  We propose a novel approach to performing merges
based on deep learning models.  Our approach depends on (a) creating
a deep learning model that maps surface forms of an entity into a
set of vectors such that alternate forms for the same entity are
closest in vector space, (b) indexing these vectors using a nearest
neighbors algorithm to find the forms that can be potentially joined
together.  To build these models, we had to adapt techniques from
metric learning, due to the properties of entity names.  We
demonstrate sample selection and loss functions, which work for this
domain.  To evaluate our approach, we used Wikidata as ground truth
and built models for datasets with approximately 1.1M people's names
(200K identities) and 130K company names (70K identities).  We found
that vector embeddings allow for joins with precision of .91-.93 and
recall of .70-.81.  We make the models available to the community
for aligning people or companies across multiple datasets.  
\end{abstract}
